---
title: "GNSS/GPS precision and error"
author: "Sylvain Klein & Cyrille Médard de Chardon"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GNSS precision & error

A Global Navigation Satellite System (GNSS), also commonly called a Global Positioning System (GPS) which is actually a brand of GNSS, are commonly believed to provide the best quality georeferencing of points. Georeferencing is the process of associating points on the surface of the Earth to a standardized coordinate system. This allows us to say the Eiffel Tower is located at 29.9792° N, 31.1342° E using the WGS 84 geographic coordinate system. This is typically just referred to as a latitude and longitude without any thought to what geographic coordinate system we are using. While coordinates can be [infinitely precise](https://xkcd.com/2170/) the question this resource aims to explore is how *accurate* are the devices that we use to record the coordinates.

A brief reminder of the difference between accuracy and precision:

| Value | Measurement | Definition |
|---|---|---|
|2.017|2.8325|Precise but not accurate|
|2.017|2|Accurate but not precise|
|2.017|2.017|Accurate and precise|
Note that it is better to be accurate rather than precise.

## Actual path
Below is the path taken by Sylvain from a parking garage at the train station to the MSH building in [Belval](https://www.luxatlas.lu/#b=2362&o=null&f=1071&ob=100&oo=100&of=50&x=662203&y=6360423&z=16.541415223113926). The path goes through a pedestrian area surrounded by tall structures and buildings.
```{r actual path}
# create static image
```

## Sensed path - GNSS data
Describe error over space-time

## 3D visualization - what causes error

## HDOP metric vs sat count vs ground-truthed error

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
